### 2011-05-06 draft2.tex
Introducting data types
* Grammar updated to define data type
* D[[ ]] translation for data type, selectors...

\documentclass{article}
% Test comment

@@

\newcommand{\etrans}[1]{\mathcal{E} \llbracket #1 \rrbracket}
\newcommand{\dtrans}[1]{\mathcal{D} \llbracket #1 \rrbracket}
\newcommand{\strans}[1]{\mathcal{S} \llbracket #1 \rrbracket}
\newcommand{\trans}[1]{\llbracket #1 \rrbracket}

@@

\subsection{$\lambda$-lifted haskell subset}
\begin{center}
\begin{array}{rclr}
  u,e &:=& x \mid f \mid e~e \mid D \mid \bad \mid n & (Expression)\\
  p &:=& \Delta_1 \dots \Delta_n & (Program)\\
  \Delta &:=& d \mid transp (d,c) \mid opaque (d,c) & (Defintion~and~contract)\\
  d &:=& f~x_1 \dots x_n = e \mid f~x_1 \dots x_n = \mbox{ case } e \mbox{ of } [(pat_i,e_i)] & (Definition)\\
  pat &:=&D~x_1 \dots x_n & (Pattern)\\
\end{array}
\end{center}

\subsection{FOL}
\begin{center}
\begin{array}{rclr}
  t &:=& x \mid \mbox{app}(t_1,t_2) \mid D \mid f \mid n \mid \bad \mid \unr \mid \mbox{CF}(t) & (Term)\\
  \phi &:=& \forall x.\phi \mid \phi \to \phi \mid \lnot \phi \mid \phi \lor \phi \mid \phi \land \phi \mid true \mid t=t \mid \mbox{CF}(t)& (Formula)\\
\end{array}
\end{center}
CF$(t)$ holds iff $t$ satisfies $\ok$.

@@

\section{Translation}
We define several translations: $\etrans{},\dtrans{},\strans{},\trans{}$.
\begin{center}
\begin{array}{rclr}
  \etrans{} &::& Expression \to Term\\
  \dtrans{} &::& Definition \to FOF\\
  \strans{} &::& Expression \to Contract \to FOF\\
  \trans {} &::& Definition \to Contract \to FOF\\
\end{array}
\end{center}

@@

\begin{eqnarray}
\etrans{x} &=& x\\
\etrans{f} &=& f\\
\etrans{e_1~e_2} &=& app(e_1,e_2)\\
\etrans{D} &=& D\\
\etrans{\bad} &=& \bad\\
\etrans{n} &=& n\\

@@

\begin{eqnarray}
\dtrans{f~x_1 \dots x_n = e} &=& \forall x_1 \dots x_n.\etrans{f~x_1 \dots x_n} = \etrans{e}\\
\dtrans{f~x_1 \dots x_n = \mbox{case } e \mbox{ of } [D_i~\overline{z} \mapsto e_i]} &=& \forall x_1 \dots x_n~(\bigwedge_i (\forall \overline{z}~\etrans{e} = \etrans{D_i~\overline{z}} \to \etrans{f~x_1 \dots x_n} = \etrans{e_i})\\
&& \land \etrans{e} = \bad \to \etrans{f~x_1 \dots x_n} = \bad)\\ 
&& \land \etrans{f~x_1 \dots x_n} = \unr \bigvee_i (\texttt{HD}(e) = D_i) \lor e = \bad
\end{eqnarray}

@@

\subsection{$\trans{}$}
It's the final translation, which takes a function definition and its contract and returns a first-order formula\\

@@

\begin{eqnarray}
\trans{opaque(f~x_1 \dots x_n = e,c)} &=& \dtrans{f~x_1 \dots x_n = e[f_p/f]} \land \strans{f \in c} \land \strans{f_p \in c}\\
\trans{transp(f~x_1 \dots x_n = e,c)} &=& \dtrans{f~x_1 \dots x_n = e[f_p/f]} \land \strans{f \in c} \land \strans{f_p \in c}\\
\trans{f~x_1 \dots x_n = e} &=& \trans{opaque(f~x_1\dots x_n) = e,Ok \to \dots \to Ok)}
\end{eqnarray}

We'd like a typical contract-checking session to go like this:
\begin{enumerate}
\item Start with an empty theory $T$.
\item Let $f~x_1 \dots x_n = e \in c$ be an opaque function definition to check wrt contract $c$. Check (with equinox) the consistency of the theory $T' = T \cup \trans{f~x_1 \dots x_n = e \in c}$
\item If $T'$ is consistent then let $T = \strans{f \in c} \cup T$ and go to 2. with the next function definition; otherwise give a counter-example and ask the user for refinement of the contracts and/or lemmas(?)
\end{enumerate}

\section{Questions}

Here are some open issues, design choices and equinox-related questions.

\begin{enumerate}
\item Nested implications may lead to existential quantification, is it troublesome in equinox? (altough we don't have this case here)
\item More generally, does equinox accept any FOF the grammar here defines?
\item In the section $\dtrans{}$, we believe replacing $\varphi$ by $\lor \trans{f x_1 \dots x_n} = \unr$ is equivalent. What's better for equinox?
\item Is the session stuff realistic? It looks like it can have a quadratic behaviour but maybe with the right API it's ok?
\end{enumerate}

>>>>

\documentclass{article}

@@

\newcommand{\etrans}[1]{\mathcal{E} \llbracket #1 \rrbracket}
\newcommand{\dtrans}[1]{\mathcal{D} \llbracket #1 \rrbracket}
\newcommand{\ttrans}[1]{\mathcal{T} \llbracket #1 \rrbracket}
\newcommand{\strans}[1]{\mathcal{S} \llbracket #1 \rrbracket}
\newcommand{\cf}[1]{\mbox{CF}(#1)}
\newcommand{\trans}[1]{\llbracket #1 \rrbracket}

@@

\subsection{$\lambda$-lifted haskell subset}
\begin{center}
\begin{array}{rclr}
  u,e &:=& x \mid f \mid e~e \mid D(e,\dots,e) \mid \bad \mid n & (Expression)\\
  p &:=& \Delta,p \mid T,p \mid \epsilon & (Program)\\
  %\Delta &:=& d \mid transp (d,c) \mid opaque (d,c) & (Defintion~and~contract)\\
  d &:=& f~x_1 \dots x_n = e \mid f~x_1 \dots x_n = \mbox{ case } e \mbox{ of } [(pat_i,e_i)] & (Definition)\\
  T &:=& \mbox{data } x~\tau_1 \dots \tau_n = D_1 ; D_2 ; \dots ; D_n & (Data~type~definition)\\ %TODO make type clearer
  pat &:=& D(x_1,\dots,x_n) & (Pattern)\\
\end{array}
\end{center}

\subsection{FOL}
\begin{center}
\begin{array}{rclr}
  t &:=& x \mid \mbox{app}(t_1,t_2) \mid D(t,\dots,t) \mid f \mid n \mid \bad \mid \unr \mid \mbox{CF}(t) & (Term)\\
  \phi &:=& \forall x.\phi \mid \phi \to \phi \mid \lnot \phi \mid \phi \lor \phi \mid \phi \land \phi \mid true \mid t=t \mid \mbox{CF}(t)& (Formula)\\
\end{array}
\end{center}
CF$(t)$ holds iff $t$ satisfies $\ok$.

@@

\section{Translation}
We define several translations: $\etrans{},\dtrans{},\ttrans{},\strans{}$.
\begin{center}
\begin{array}{rclr}
  \etrans{} &::& Expression \to Term\\
  \dtrans{} &::& Definition \to FOF\\
  \ttrans{} &::& Data~type  \to FOF\\
  \strans{} &::& Expression \to Contract \to FOF\\
%  \trans {} &::& Definition \to Contract \to FOF\\
\end{array}
\end{center}

@@

\begin{eqnarray}
\etrans{x} &=& x\\
\etrans{f} &=& f\\
\etrans{e_1~e_2} &=& app(e_1,e_2)\\
\etrans{D(e_1,\dots,e_n)} &=& D(\etrans{e_1},\dots,\etrans{e_n}\\
\etrans{\bad} &=& \bad\\
\etrans{n} &=& n\\

@@

\begin{eqnarray}
\dtrans{f~x_1 \dots x_n = e} &=& \forall x_1 \dots x_n.\etrans{f~x_1 \dots x_n} = \etrans{e}\\
\dtrans{f~x_1 \dots x_n = \mbox{case } e \mbox{ of } [D_i~\overline{z} \mapsto e_i]} &=& \forall x_1 \dots x_n~(\bigwedge_i (\forall \overline{z}~\etrans{e} = \etrans{D_i~\overline{z}} \to \etrans{f~x_1 \dots x_n} = \etrans{e_i})\\
&& \land \etrans{e} = \bad \to \etrans{f~x_1 \dots x_n} = \bad)\\ 
&& \land (\etrans{f~x_1 \dots x_n} = \unr \bigvee_i (\texttt{HD}(e) = D_i) \lor e = \bad)
\end{eqnarray}

@@

@@

\section{Datatypes}
$\ttrans{T}$ is a set of first-order formulae.
$\ttrans{\mbox{data } T = D_1 ; \dots ; D_n} = S_1 \cup S_2 \cup S_3$.

First, for each $D_i$ we introduce selectors $sel_{k,D_i}$, which are basically projections of $D_i(x_1,\dots,x_n)$ on its $k$-th composant, so that we can express that constructors are injective :
$$S_1 := \{ \forall x_1,\dots,x_n . \bigwedge_{1 \leq i \leq n} sel_{i,D_i}(D_i(x_1,\dots,x_n)) = x_i \mid 1 \leq i \ n \}$$

For each pair of different constructors $D_i,D_j$, we state that they can never map to the same value: 
$$S_2 := \{ \forall x_1,\dots,x_n~\forall y_1,\dots,y_k . D_i(x_1,\dots,x_n) \neq D_j(y_1,\dots,y_k) \mid 1 \leq i < j \leq n \}$$

Finally, we have to give crash-freeness conditions for each $D_i$:
$$S_3 := \{ \forall x_1,\dots,x_n . (\cf{x_1} \land \dots \land \cf{x_n} \leftrightarrow \cf{D_i(x_1,\dots,x_n)}) \mid 1 \leq i \leq n \}$$
